---
title: "0012_the_predicted_network"
author: "Dan MacLean"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(tidygraph)
library(ggraph)
library(igraph)
```

## Loading the predicted network

The prediction of TF/Target pairs was done outside the markdown using the script `scripts/run_tf_predict.R`. 

```{bash eval=FALSE}
Rscript scripts/run_tf_predict.R
```

This produced the file "predictions.csv", which has the columns 

```{r echo = FALSE}
knitr::kable(data.frame(TF_AGI = NA, TARGET_AGI = NA, TF_affy = NA, TARGET_affy = NA, prediction = NA, likelihood = NA))
```

I'll load the predictions with likelihood >= 0.99 into a dataframe and process into a graph. I'll reorder to use the affy names as id, since they're unique.

```{r}
interactions_df <- readr::read_csv("../data/predictions.csv", 
                col_names = c("TF_AGI", "Target_AGI", 
                              "TF_affy", "Target_affy", 
                              "prediction_class", "likelihood"
                              )
                ) %>% 
  filter(likelihood >= 0.99) %>%
  select(TF_affy, Target_affy, TF_AGI, Target_AGI, prediction_class, likelihood)



g <- as_tbl_graph(interactions_df, directed = TRUE, mode = "out")

```

## Large-scale structure

### Components
```{r}
count_components(g)
```

### Degree distributions

```{r}

out_d <- degree(g,  mode = "out")

table(out_d)

out_degree_df <- as.data.frame(table(out_d)) %>%
 dplyr::rename(out_degree = out_d) %>%
 dplyr::mutate(out_degree = as.numeric(as.character(out_degree)))

ggplot(out_degree_df) + aes(x = log10(out_degree + 1), y = Freq) + geom_point() +
  scale_y_log10() + 
  annotation_logticks(sides = "bl") + theme_bw()

in_d <- degree(g,  mode = "in")

table(in_d)
in_degree_df <- as.data.frame(table(in_d)) %>%
  dplyr::rename(in_degree = in_d) %>%
  dplyr::mutate(in_degree = as.numeric(as.character(in_degree)))

ggplot(in_degree_df) + aes(x = log10(in_degree + 1), y = Freq) + geom_point() +
#   breaks = scales::trans_breaks("log10", function(x) 10^x),
#  scale_y_continuous("Frequency\n(how many of them)",
#   labels = scales::trans_format("log10", scales::math_format(10^.x))) + geom_smooth(method = "lm") + 
#  scale_x_log10() + 
  scale_y_log10() + 
  annotation_logticks(sides = "bl") + theme_bw()

```

### In/Out Structure

Divide nodes into TF/Regulator and with in/out, in only (sink), out only (source)

```{r}
tfs <- unique(interactions_df$TF_affy)
targets <- unique(interactions_df$Target_affy)

#V(g)$is_tf <- V(g)$name %in% tfs
#V(g)$is_target <- V(g)$name %in% targets

g <- g %>% activate(nodes) %>%
  mutate(is_tf = name %in% tfs) %>%
  mutate(out_degree = degree(g, mode = "out") ) %>%
  mutate(in_degree = degree(g, mode = "in") ) %>% 
  mutate(type = if_else(out_degree == 0, "sink", if_else(in_degree == 0, "source", "link"))) 

ggraph(g, "hive", axis = "type", sort.by = "out_degree") + 
  geom_edge_hive(alpha = 0.1) +
  geom_axis_hive(aes(colour = type), label = FALSE) + coord_fixed() + theme_bw()

```

Looks like the majority of the edges here are source to sink, direct. Perhaps not much structure, in the network overall. Only a few of the very highest out degree sources pass to a link then presumably to a source. I'll look at the degree distributions of the types of node, then remove the source -> sink direct edges.


```{r}
g %>% activate(nodes) %>%
  as_tibble() %>%
  ggplot() + aes(out_degree) + 
  geom_density(aes(colour= type, fill = type)) +
  facet_wrap( ~ type, scales = "free") 

g %>% activate(nodes) %>%
  as_tibble() %>%
  ggplot() + aes(in_degree) + 
  geom_density(aes(colour= type, fill = type)) +
  facet_wrap( ~ type, scales = "free") 
```

Actually, Im missing a step. This is really the wrong network to be doing this on. I need to work out which treatments the activation is happening in and analyse sub networks. Back to the drawing board!